---
title: Regression problem
subtitle: The release year prediction of the songs from the timbre features.
author: "Mateusz Baryła & Agata Wytrykowska"
bibliography: bibliography.bib
link-citations: yes
nocite: '@*'
output:
html_document:
    toc: true
    toc_float: true
    number_sections: true
    theme: united
    df_print: paged
---

# Description of data and problem analyzed

Audio features play a significant role in the prediction of the song release year. The Million Song Dataset is a free collection of million music tracks consists. Based on this dataset a subset has been created that returned mostly the songs between 1922 to 2011. Most of the observations relate to the  2000s [@Bertin-Mahieux2011]. We downloaded it from the Machine Learning Repository [@Dua:2019].

```{r, echo = FALSE}
pacman::p_load(data.table, tidyverse, knitr,
               GGally, ggcorrplot, tidymodels,
               randomForest, Rcpp, ranger,
               kableExtra, tree, rpart, caret,
               VSURF)
experiment <- FALSE
```

There are 515345 observations in the dataset. 90 attributes describe each and single observation. All of the attributes are real number and correspond to the timbre either average or covariance. The Echo Nest API is used in order to extract the features from the timbre which must be an additional service. In overall, timbre-related features describes spectogram patches. Considering the particular dimensions of the 12-dimension timbre vector they represent the average loudness of the segment, with the second the brightness is described, the third relate to the flatness of a sound and the fourth to the sounds with stronger attack [@Bertin-Mahieux2013].

## Data Preprocessing

Data preprocessing is told to take around 80 % of the data science project. In our case the data is very well prepared, there

## Reading data
```{r, echo = FALSE}
source('..\\src\\utils\\fit_table.r')
```

```{r}
if (experiment){
  raw_data <- read.csv2('..\\data\\raw\\YearPredictionMSD_sample.txt')
} else{
  raw_data <- read.csv('..\\data\\raw\\YearPredictionMSD.txt', header = FALSE)
}
raw_data %>% fit_table()
```

The first column in the dataset correspond to the year that will be predicted from the features. With the MSD many tasks can be addressed, however the song release year brings many practical applications and there are not a lot of research papers that relate to this prediction problem. Only audio features are used in this problem which is justified with the fact that listeners are often fond of music from certain periods of their lives. Another reason might be the evolution of timbre across years.

## Renaming columns

We rename the first column for the sake of the modelling purposes. 

```{r}
raw_data <- raw_data %>%
        rename(year = V1) %>%
        mutate(year = as.numeric(year))
raw_data %>% fit_table()
```

## Scaling

The authors in the main article mapped the year values using linear mapping to [0,1] range [@Bertin-Mahieux2011]. We follow the convention in the analysis below.

```{r}
# mapping target to [0,1] range
range_to_0_1 <- function(vec) {
  (vec - min(vec)) / (max(vec) - min(vec))
}

raw_data$year_scaled <- range_to_0_1(raw_data$year)
raw_data$year_scaled <- range_to_0_1(raw_data$year)
```

## Splitting covariates

```{r}
timbre_average_data <- raw_data %>%
        select(V2:V13)
timbre_average_data %>% fit_table()
```

We needed to investigate what the timbre in the sound analysis actually means. It is described as everything that can be differentiated between two different tracks by the listener having kept the same pitch, spatial location and loudness and refers to the perceptual quality of a sound and color. "Brightness", "clarity", "harshness", "fullness" and "noisiness" are frequently used adjectives that are associated with the timbre word [@Allen2018].

```{r}
timbre_covariance_data <- raw_data %>%
        select(V14:V91)
timbre_covariance_data %>% fit_table()
```

Covariance is taken over all segments, where each segment is described by 12-dimensional vector.

## Handling missing values

```{r}
raw_data %>%
        summarise(across(everything(), ~sum(is.na(.))))
```
There are no missing values across columns and in the whole dataset. This is an important insight. It is likely to influence an easier later phases of the project e.g exploratory data analysis and modelling.

# Descriptive analyses of the data

Having prepared the dataset we can follow the analysis and in the next points visualized dataset including both target and features on which the model will be trained.

## One-dimensional plots

The target variable in the problem analyzed is year. We easily spot the distribution of it being heavily skewed.

```{r}
raw_data %>%
    ggplot(aes(x = year)) +
    geom_histogram(aes(y = ..density..),
                   bins = 50, alpha = 0.4, fill = "pink", col = "gray30") +
    geom_density(alpha = .2, fill = "blueviolet") +
    labs(title = "Distribution of the outcome variable - density and histogram",
         caption = "source: produced with ggplot package") +
    theme(plot.title = element_text(hjust = 0.5, size = 20),
          axis.text.x = element_text(size = 20))
```

## Two-dimensional plots

### Number of songs across centuries

For the first plot we have decided to squeeze the year variable based on the century because the distribution of the songs across years is highly nonuniform.

```{r}
timbre_average_data %>%
    mutate(year = raw_data$year) %>%
    mutate(century = if_else(year <= 2000, 'XX', 'XXI')) %>%
    group_by(century) %>%
    count() %>%
    ggplot(aes(x = century, y = n)) +
    geom_bar(stat = "identity") +
    ggtitle('The songs count across centuries') +
    labs(caption = "source: produced with ggplot package",
         x = 'Century',
         y = 'Count') +
    theme_minimal() +
    theme(legend.position = 'bottom')  +
    guides(fill = guide_legend(nrow = 2,byrow = TRUE)) +
    scale_fill_grey() +
    theme(plot.title = element_text(hjust = 0.5, size = 20))
```

We see that the count of songs increase with the centuries, although, there are more years in the dataset, in XX century.

### Loudness across decades

```{r}
timbre_average_data %>%
    mutate(year = raw_data %>% pull(year)) %>%
    mutate(decade = as.factor(floor(year/10)*10)) %>%
    rename(loudness = V2) %>%
    select(loudness, decade) %>%
    ggplot(aes(x= decade, y = loudness, group=decade, fill = decade)) +
    geom_boxplot() +
    labs(title = "Loudness across decades") +
    guides(fill = "none", group = "none") +
    theme(plot.title = element_text(hjust = 0.5))
```

Increase in loudness is spotted across decades. Nowadays people listen to louder music more.

### Correlation matrix

An important area that should not be neglected in the process of building machine learning model is creating the correlation matrix based on which we will continue investigation regarding the two-dimensional plots. We have decided to learn more about the timbre split suggested by the authors. Based on that we can check whether average timbre or timbre covariance have stronger influence on the outcome variable.

```{r}
corr_average <- timbre_average_data %>%
        mutate(year = raw_data$year) %>%
        cor()

ggcorrplot(corr_average, hc.order = TRUE, type = "lower", lab = TRUE) +
    labs(title = "Correlation between timbre average data and the outcome variable.") +
    theme(plot.title = element_text(hjust = 0.5, size = 20))
```

Based on the correlation matrix we spot two variables that might be important in predicting the song release year V2 and V7. V7 is negatively correlated with the year which means that the higher the variable the lower the year.

```{r}
corr_covariance <- timbre_covariance_data %>%
        mutate(year = raw_data$year) %>%
        cor()
```

```{r}
high_corr_columns <- corr_covariance %>%
        as_tibble() %>%
        mutate(row_name = c(timbre_covariance_data %>% names(), "year")) %>%
        relocate(row_name) %>%
        pivot_longer(-row_name) %>%
        filter(name == 'year') %>%
        filter(value > 0.05) %>%
        pull(row_name)
```


Because there are a lot of columns inside the timbre covariance matrix we choose to visualize only the ones that are greater than chosen level of correlation. Interestingly it is set very low to 0.05 since the association is extremely low.

```{r}
corr_covariance <-  timbre_covariance_data %>%
        mutate(year = raw_data$year) %>%
        select(all_of(high_corr_columns)) %>%
        cor()
```

```{r}
ggcorrplot(corr_covariance,
           hc.order = TRUE,
           type = "lower",
           lab = TRUE,
           method = 'circle',
           lab_size = 3) +
  labs(title = "Correlation between timbre covariance data and the outcome variable.") +
  theme(plot.title = element_text(hjust = 0.5, size = 20))
```

The highest correlation with the target variable is spotted between variable V21, V74, V15 and V40 What is more the multicollinearity might be spotted for variable V15 and V21 with the correlation value 0.48.

## Statistical analysis

```{r}
source("..\\src\\utils\\t_test_across_centuries.R")
t_test_test_across_centuries(timbre_average_data %>%
                                     mutate(year = target_data$year),
                             'V2')
t_test_test_across_centuries(timbre_average_data %>%
                                     mutate(year = target_data$year),
                             'V7')
t_test_test_across_centuries(timbre_covariance_data %>%
                                     mutate(year = target_data$year),
                             'V21')
t_test_test_across_centuries(timbre_covariance_data %>%
                                     mutate(year = target_data$year),
                             'V15')
test_test_across_centuries(timbre_covariance_data %>%
                                     mutate(year = target_data$year),
                             'V74')
```

We have show that across the centuries averages differ for the variables that are the most correlated with the target variable. It might be an indicator that these variables will play a significant role in predicting the song release year.

# Variable transformation

In this step we embrace creating interactions between the most influential variables accordingly to the descriptive analysis step.

## Interactions

Since we deal with only numeric variables we do not need to perform dummification step.

```{r}
formula_for_interactions <- year_scaled ~ V2 + V7 + V21 + V15 + V74

rec <- recipe(formula_for_interactions, data = raw_data ) # is it correct?

interact_terms <- rec %>%
  step_interact(terms = ~ starts_with("V"):starts_with("V"))

interact_prep <- prep(interact_terms, training = raw_data)

transformed_data <- bake(interact_prep, raw_data)

transformed_data %>% names()
```

Final transformed data used in the modelling consist of 16 dimensions. They are a result of variable transformation step.

# Variable selection

We investigated two packages varSelRF [@varselrf] and VSURF [@genuer:hal-01251924] for variable selection step. We have chosen them mainly because of their dependency on the randomForest package that was used during the classes. The latter one is not implemented for regression problem. Hence, varSelfRF will be used which is more complicated procedure. It consists of three steps: thresholding, interpretation and prediction step. Thresholding is based on variable importance. The interpretation embeds random forest models starting with the one built only on the variables chosen in the previous step. Next, the stepwise selection procedure is applied in the final, prediction procedure.

The result of this step already trains the first model. We will evaluate its performance during the comparison of the trained machine learning models.

# Training/test data division and resampling

## Splitting dataset

The result might suggest that the dataset is somehow not completely random. It is confirmed with the fact that the following train / test split should be respected. First 463 715 observations should be used for the training purposes while the remaining 51 630 are the observations on which the regression model will be tested.

The dataset authors add that thanks to that split 'producer effect' will be avoided. Next, we confirm that the remaining attributes correspond to the timbre covariates.


```{r}
if (experiment){
  train_data <- transformed_data %>% slice_head(n = 890)
  test_data <- transformed_data %>% slice_tail(n = 110)
} else{
  train_data <- transformed_data %>% slice_head(n = 463715)
  test_data <- transformed_data %>% slice_tail(n = 51630)
}
```

We considered all the variables in the dataset, however the problem appear to be unfeasible. We experienced both memory and time complexity. We try to omit the obstacles with the large size of the dataset. We choose only these dimensions shown in descriptive analysis that are correlated with the target above the manually specified threshold.

# Different machine learning methods

## Basic tree model

Therefore, we operate on the transformed dataset with interactions considering only chosen variables.

```{r}
model_formula <- year_scaled ~ .
```

We start with the basic tree model - `tree()`. It uses the *binary recursive algorithm*. Splits are generated until the criterium that relates to node impurity is not met.

```{r}
basic_tree <- tree(model_formula, train_data)
```

```{r}
summary(basic_tree)
```

As presented above the model chosen only two variables for the final prediction. Both of them include loudness.

```{r}
plot(basic_tree)
text(basic_tree, pretty = 0)
```

Interestingly, there are only three terminal nodes. It might relate to the skewness of the target variable. Nevertheless,
it can be somehow described as naive, baseline model.

## Tree model with rpart

Package rpart offer different way of handling surrogate variables.

```{r}
rpart_tree <- rpart(model_formula,
                     data = train_data,
                     method = "anova")
```

### optimization

We want to optimize  `minsplit`. It relates to the number of observations that need to belong to the node before the split is considered. We neglect `maxdepth` because of already contrained number of final variables included in the model.

```{r, results='hide'}
hyper_vec <- expand.grid(
        minsplit = seq(2, 8, 1)
)
```
The values in the grid are chosen so that they relate to the data that we are using in this problem.

```{r, results='hide'}
vec_models <- list()
for (comb in 1:nrow(hyper_vec)) {

  cat(comb, "/", nrow(hyper_vec), "\n", sep = "")

  minsplit <- hyper_vec$minsplit[comb]

  # settin the seed
  set.seed(9438 + comb)

  # training of the model and saving results to the list
  vec_models[[comb]] <- rpart(
          formula = model_formula,
          data    = train_data,
          method  = "anova",
          control = list(minsplit = minsplit)
  )
}
```

Now models with the lowest error can be easily chosen. Therefore,
we use the functions that were introduced during the classes.

```{r }
get_cp <- function(x) {
  min <- which.min(x$cptable[, "xerror"])
  cp  <- x$cptable[min, "CP"]
  return(cp)
}

get_min_error <- function(x) {
  min    <- which.min(x$cptable[, "xerror"])
  xerror <- x$cptable[min, "xerror"]
  return(xerror)
}
```

Last but not least, we choose top three performers out 7 different models.

```{r, echo = FALSE}
hyper_vec %>%
        mutate(
                cp    = purrr::map_dbl(vec_models, get_cp),
                error = purrr::map_dbl(vec_models, get_min_error)
        ) %>%
        arrange(error) %>%
        top_n(-3, wt = error)
```

Let's build tree with the parameters that correspond to the lowest prediction error on the testing set.

```{r }
optimized_tree <- rpart(
        formula = model_formula,
        data    = train_data,
        method  = "anova",
        control = list(minsplit = 5, cp = 0.01) #to put after experiment is completed
)
```

Optimized tree is the one with parameters that correspond to the lowest error.

## XGBoost

As the next step, we use XGBoost. What is more, we additionally center and scale the predictors.

```{r}
trainctrl <- trainControl(verboseIter = TRUE,
                          method = "cv",
                          number = 10)

xgb_model <-
  train(model_formula,
        data = train_data,
        method = "xgbTree",
        trControl = trainctrl,
        preProcess = c("center", "scale"))
```

## VSURF

Variable selection occurred to be the most challenging process. It seems to be computationally infeasible. We limit the training data to sample 10000 observations.

```{r}
data_for_vsurf <- train_data %>% sample_n(10000)
vsurf <- VSURF(x = data_for_vsurf %>% select(-year_scaled),
               y = data_for_vsurf %>%  pull(year_scaled),
               ntree = 100, nfor.thres = 13,
               parallel = 4,
               nfor.interp = 5, nfor.pred = 3)
```

# Comparison

In this step we compare models built in the report and their accuracy on the test data. We will use script from the classes - getRegressionMetrics.

```{r}
source('..\\src\\models\\getRegressionMetrics.R')
```

We add yielding the results in a form of a readable table.

```{r }
performance_comparison <- bind_cols(tibble(tree         =  predict(basic_tree,  test_data),
                 vsurf = predict(vsurf, test_data) %>% pull(pred),
                 xgb = predict(xgb_model, test_data),
                 optimized_tree         =  predict(optimized_tree,  test_data))) %>%
        map_dfr(getRegressionMetrics, real = test_data$year_scaled, .id = "model")
```

# Summary and conclusions

First of all, for the first time we have been modelling on dataset of this size.
We needed to take appropriate measures in order to experiment faster for instance
sampling 1000 observations and working only on them.

Secondly, the variables' names were not provided and we managed to find the naming for only few of them. We were not aware how the timbre might influence the outcome variable and haven't any intuition towards the variables that might be important. Hence, we focused on quantitative measures such as correlations or statistical tests. Nevertheless, the visualizations also occurred to be informative.

What is more, the variable selection process was an opportunity to learn more about
how to choose packages. During the notebook production we have already known
what libraries were introduced during the classes (randomForest), what machine learning problem we analyze (regression) and what are the methods that it needs to implement (predict). Considering these three constraints we have chosen one that suits all of the aforementioned needs.

Last but not least, it has been a fruitful experience to have all the results
stored in a single table and compare them easily. Based on them we see that how
the performance measures are distributed across models.

# References

# Appendix

```{r}
source('..\\src\\utils\\fit_rf_model_with_recipe.R')
simple_recipe_with_log <-
        recipe(year_scaled ~ .,
               data = train_data %>% mutate(year = target_data$year_scaled)) %>%
                step_log(V15, base = 10)
```


```{r}
simple_recipe_with_interact <-
        recipe(year_scaled ~ .,
               data = train_data) %>%
                step_interact(terms = ~V14:V91)
```

## Model with transformed variables
```{r}
fit_rf_model_with_recipe(simple_recipe_with_log, train_data)

fit_rf_model_with_recipe(simple_recipe_with_interact, train_data)
```