---
title: Classification problem
subtitle: Good/bad credit risk classification for people described by a set of attributes.
author: "Mateusz Baryła & Agata Wytrykowska"
bibliography: bibliography.bib
link-citations: yes
output:
  html_document:
    toc: yes
    df_print: paged
  pdf_document:
    toc: yes
    df_print: kable
    fig_width: 7
    fig_caption: yes
    html_document:
      number_sections: no
      theme: cerulean
      highligt: pygments
---

# classification



## Introduction*



## Data Inspection


## Data Preprocessing
###Loading data

### Reading data 

```{r}
####Agata - Loading Data
df <- read.csv("~/Desktop/studia/magisterka/3sem/ML2/mlpro/dataset_31_credit-g.csv")

####Mateusz - Loading Data
pacman::p_load(tidyverse)
df <- read.csv("classification\\data\\raw\\dataset_31_credit-g.csv")
```

### Hadnling missing data 

``` {r, include = TRUE}
table(is.na(df))

```

There is no missing values in the dataset. 
There is no need to verify a method of handling missing values. 

### Spotting variables' duplicates 
``` {r, include = TRUE}

table(duplicated(df))
```
There is no duplicates in data. 

### Transformation of variables 
``` {r, include = TRUE}
###transformation of variables changing characters into factors
sapply(df, table)

df_f <- as.data.frame(unclass(df), stringsAsFactors = TRUE) # Convert all character columns into factors
sapply(df_f, class)

df_f %>% as_tibble()
```



### Types, scales features of data 







## Model

## Diagnostics
<!-- Residuals Error Checking -->

## Forecast

## Analysis

## Conclusion

## References
<!-- This is required to attach the bibliography -->