---
title: Classification problem
subtitle: Good/bad credit risk classification for people described by a set of attributes.
author: "Mateusz Baryła & Agata Wytrykowska"
bibliography: bibliography.bib
link-citations: yes
output:
  html_document:
    toc: yes
    df_print: paged
  pdf_document:
    toc: yes
    df_print: kable
    fig_width: 7
    fig_caption: yes
    html_document:
      number_sections: no
      theme: cerulean
      highligt: pygments
---

# Description of data and problem analyzed
The dataset was donated in 1994. It consists of 20 attributes and 1000 instances. We downloaded it from Machine Learning repository [@Dua:2019].Two classes: "good" and "bad" credits are associated with the dataset. Within 20 variables there are hidden 13 categorical ones and 7 numerical. The dataset is unbalanced with 700 observations that belong to the "good" class and 300 belonging to the "bad" one [@Wang2003]. As stated in the literature it is a result of experiment. Each applicant was described by a set of attributes [@ODea2001].

## Initial descriptive analyses of the data

## Reading data
```{r, echo = FALSE}
source('..\\src\\utils\\fit_table.r')
```

```{r}
####Agata - Loading Data
df <- read.csv("~/Desktop/studia/magisterka/3sem/ML2/mlpro/dataset_31_credit-g.csv")
####Mateusz - Loading Data
pacman::p_load(tidyverse)
df <- read.csv("..\\data\\raw\\dataset_31_credit-g.csv")
df %>% fit_table()
```
## Renaming columns

We prepared the mapping table that will serve in the next phases of the project - for instance
during the exploratory data analysis.
```{r}
#Types of variables
features <- df %>% names()
type <- c('qualitative', 'quantitative', 'qualitative', 'qualitative',
          'quantitative', 'qualitative', 'qualitative', 'quantitative',
          'qualitative', 'qualitative', 'quantitative', 'qualitative',
          'quantitative', 'qualitative', 'qualitative', 'quantitative',
          'qualitative', 'quantitative', 'qualitative', 'qualitative',
          'qualitative')
specific_type <- c('discrete', 'discrete', 'discrete', 'discrete',
                   'continuous', 'discrete', 'discrete', 'continuous',
                   'discrete', 'discrete', 'discrete', 'discrete',
                   'discrete', 'discrete', 'discrete', 'discrete',
                   'discrete', 'discrete', 'discrete', 'discrete',
                   'discrete')
scales <- c('ordinal', 'ordinal', 'nominal', 'nominal',
            'ratio', 'ordinal', 'ordinal', 'ratio',
            'nominal', 'nominal', 'ratio', 'nominal',
            'ratio', 'nominal', 'nominal', 'ordinal',
            'ordinal', 'discrete', 'nominal', 'nominal',
            'nominal')
pretty_names <- c('Status of existing checking account', 'Duration in month',
                  'Credit history', 'Purpose of taking a credit', 'Credit amount',
                  'Savings account/bonds', 'Present employment since',
                  'Installment rate in percentage of disposable income',
                  'Personal status and sex', 'Other debtors / guarantors',
                  'Present residence since', 'Property',
                  'Age in years', 'Other installment plans',
                  'Housing', 'Number of existing credits at this bank',
                  'Job', 'Number of people being liable to provide maintenance for',
                  'Telephone', 'Foreign worker', 'Class of a customer')
dictionary <- tibble(features = features,
                     general_type = type,
                     specific_type = specific_type,
                     scale = scales,
                     pretty_name = pretty_names)
dictionary
```


With that dictionary we can much easier examine the overall nature of the dataset. Types of variables,
their scales and names in a readable format will more likely allow for the more effective
experimentation in the next phases of the project.

## Splitting dataset
```{r}
train_data <- raw_data %>% slice_head(n = 666)
test_data <- raw_data %>% slice_tail(n = 334)
```

We follow the split suggested to having 66 % observations in the train dataset
and remaining 34 % in the test dataset [ODea2001].

# Descriptive analyses of the data

## Distributions

Assessing the distribution of the credit amount across groups
might be interesting. Hypothetically, good clients are these with larger
credit amount. If they weren't classified to that group they would not
be given a credit.

```{r}
df %>%
  ggplot(aes(credit_amount, color = class)) +
  geom_density() +
  stat_density(geom = 'line', position = 'dodge', size = 1) +
  labs(title = "Comparison of credit amount among group of respondents.") +
  theme(plot.title = element_text(hjust = 0.5))
```

## Two-dimensional plots

## Variable transformations

## Variable selection methods

## Training/test data division, resampling (e.g. cross-validation, down-sampling, up-sampling)

## Comparison of prediction accuracy of at least 3 different machine learning methods on test data

## Summary and conclusions

### Hadlding missing data

```{r, include = TRUE}
table(is.na(df))
```

There are no missing values in the dataset.
There is no need to verify a method of handling missing values. 

### Spotting variables' duplicates 
```{r, include = TRUE}
table(duplicated(df))
```
There are no duplicates in data.

### Transformation of variables 
```{r, include = TRUE}
###transformation of variables changing characters into factors
sapply(df, table)

df_f <- as.data.frame(unclass(df), stringsAsFactors = TRUE) # Convert all character columns into factors
sapply(df_f, class)

df_f %>% as_tibble()
```

### Types, scales features of data

```{r}
#Types of variables
features <- df %>% names()
type <- c('qualitative', 'quantitative', 'qualitative', 'qualitative',
          'quantitative', 'qualitative', 'qualitative', 'quantitative',
          'qualitative', 'qualitative', 'quantitative', 'qualitative',
          'quantitative', 'qualitative', 'qualitative', 'quantitative',
          'qualitative', 'quantitative', 'qualitative', 'qualitative',
          'qualitative')
specific_type <- c('discrete', 'discrete', 'discrete', 'discrete',
                   'continuous', 'discrete', 'discrete', 'continuous',
                   'discrete', 'discrete', 'discrete', 'discrete',
                   'discrete', 'discrete', 'discrete', 'discrete',
                   'discrete', 'discrete', 'discrete', 'discrete',
                   'discrete')
scales <- c('ordinal', 'ordinal', 'nominal', 'nominal',
            'ratio', 'ordinal', 'ordinal', 'ratio',
            'nominal', 'nominal', 'ratio', 'nominal',
            'ratio', 'nominal', 'nominal', 'ordinal',
            'ordinal', 'discrete', 'nominal', 'nominal',
            'nominal')
pretty_names <- c('Status of existing checking account', 'Duration in month',
                  'Credit history', 'Purpose of taking a credit', 'Credit amount',
                  'Savings account/bonds', 'Present employment since',
                  'Installment rate in percentage of disposable income',
                  'Personal status and sex', 'Other debtors / guarantors',
                  'Present residence since', 'Property',
                  'Age in years', 'Other installment plans',
                  'Housing', 'Number of existing credits at this bank',
                  'Job', 'Number of people being liable to provide maintenance for',
                  'Telephone', 'Foreign worker', 'Class of a customer')

dictionary <- tibble(features = features,
                     general_type = type,
                     specific_type = specific_type,
                     scale = scales,
                     pretty_name = pretty_names)
dictionary
```
With that dictionary we can much easier examine the overall nature of the dataset. Types of variables, their scales and names in a readable format will more likely allow for the more effective experimentation in the next phases of the project.

## Model

## Diagnostics
<!-- Residuals Error Checking -->

## Forecast

## Analysis

## Conclusion

## References