---
title: Classification problem
subtitle: Good/bad credit risk classification for people described by a set of attributes.
author: "Mateusz Baryła & Agata Wytrykowska"
bibliography: bibliography.bib
link-citations: yes
output:
  html_document:
    toc: yes
    df_print: paged
  pdf_document:
    toc: yes
    df_print: kable
    fig_width: 7
    fig_caption: yes
    html_document:
      number_sections: no
      theme: cerulean
      highligt: pygments
---

# classification



## Introduction*



## Data Inspection


## Data Preprocessing
###Loading data

### Reading data 

```{r}
####Agata - Loading Data
df <- read.csv("~/Desktop/studia/magisterka/3sem/ML2/mlpro/dataset_31_credit-g.csv")

####Mateusz - Loading Data
pacman::p_load(tidyverse)
df <- read.csv("classification\\data\\raw\\dataset_31_credit-g.csv")
```

### Hadnling missing data 

``` {r, include = TRUE}
table(is.na(df))

```

There is no missing values in the dataset. 
There is no need to verify a method of handling missing values. 

### Spotting variables' duplicates 
``` {r, include = TRUE}

table(duplicated(df))
```
There is no duplicates in data. 

### Transformation of variables 
``` {r, include = TRUE}
###transformation of variables changing characters into factors
sapply(df, table)

df_f <- as.data.frame(unclass(df), stringsAsFactors = TRUE) # Convert all character columns into factors
sapply(df_f, class)

df_f %>% as_tibble()
```



### Types, scales features of data 

```{r}
#Types of variables
features <- df %>% names()
type <- c('qualitative', 'quantitative', 'qualitative', 'qualitative',
          'quantitative', 'qualitative', 'qualitative', 'quantitative',
          'qualitative', 'qualitative', 'quantitative', 'qualitative',
          'quantitative', 'qualitative', 'qualitative', 'quantitative',
          'qualitative', 'quantitative', 'qualitative', 'qualitative',
          'qualitative')
specific_type <- c('discrete', 'discrete', 'discrete', 'discrete',
                   'continuous', 'discrete', 'discrete', 'continuous',
                   'discrete', 'discrete', 'discrete', 'discrete',
                   'discrete', 'discrete', 'discrete', 'discrete',
                   'discrete', 'discrete', 'discrete', 'discrete',
                   'discrete')
scales <- c('ordinal', 'ordinal', 'nominal', 'nominal',
            'ratio', 'ordinal', 'ordinal', 'ratio',
            'nominal', 'nominal', 'ratio', 'nominal',
            'ratio', 'nominal', 'nominal', 'ordinal',
            'ordinal', 'discrete', 'nominal', 'nominal',
            'nominal')
pretty_names <- c('Status of existing checking account', 'Duration in month',
                  'Credit history', 'Purpose of taking a credit', 'Credit amount',
                  'Savings account/bonds', 'Present employment since',
                  'Installment rate in percentage of disposable income',
                  'Personal status and sex', 'Other debtors / guarantors',
                  'Present residence since', 'Property',
                  'Age in years', 'Other installment plans',
                  'Housing', 'Number of existing credits at this bank',
                  'Job', 'Number of people being liable to provide maintenance for',
                  'Telephone', 'Foreign worker', 'Class of a customer')

dictionary <- tibble(features = features,
                     general_type = type,
                     specific_type = specific_type,
                     scale = scales,
                     pretty_name = pretty_names)
dictionary
```
With that dictionary we can much easier examine the overall nature of the dataset. Types of variables, their scales and names in a readable format will more likely allow for the more effective experimentation in the next phases of the project.

## Model

## Diagnostics
<!-- Residuals Error Checking -->

## Forecast

## Analysis

## Conclusion

## References
<!-- This is required to attach the bibliography -->